---
title: "Estimating Heterogeneous Effects"
subtitle: "Data Analytics and Visualization with R<br>Session 8"
title-slide-attributes:
  data-background-size: stretch
  data-slide-number: none
auto-stretch: false
institute: "University of Mannheim<br>Spring 2023"
author: "Viktoriia Semenova"
footer: "[🔗 r4da.live](https://r4da.live/)"
logo: images/logo.png
format:
  revealjs:
    theme: ../slides.scss
    transition: fade
    incremental: true   
    slide-number: true
    chalkboard: true
execute:
  echo: true
editor_options: 
  chunk_output_type: console
editor: 
  markdown: 
    wrap: 80
---

```{r setup, include=FALSE}
p_needed <- c("tidyverse", "janitor", "icons", "infer",
              "countdown", "showtext", "ggdag", "magrittr",
              "gt", "plotly", "broom", "patchwork")

# check if they are already installed, install if not installed
lapply(p_needed[!(p_needed %in% rownames(installed.packages()))], install.packages, repos = "http://cran.us.r-project.org")

# load the packages
lapply(p_needed, library, character.only = TRUE)

# set width of code output
options(width = 90)

# set figure parameters for knitr
knitr::opts_chunk$set(
  fig.width = 7, # 7" width
  fig.asp = 0.618, # the golden ratio
  fig.retina = 3, # dpi multiplier for displaying HTML output on retina
  fig.align = "center", # center align figures
  dpi = 300 # higher dpi, sharper image
)


font_add_google(name = "Gochi Hand")
showtext::showtext_auto()
# set theme for ggplot2
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 14)) +
  theme(plot.title = element_text(face = "bold"))

# governors <- readr::read_csv("https://r4da.live/files/data/external_data/governors.csv")
# trains <- readr::read_tsv("../data/trains.tsv") %>% clean_names()
evals <- read_csv("../data/beauty.csv") %>% clean_names()
set.seed(2508)
```

# Intro

## Quiz: Which of these statements are correct? {.smaller}

```{r}
#| echo: false
countdown(minutes = 4, color_background = "white", left = 0)
```


::: columns
::: {.column width="50%"}
1. $R^2$ is an important criterion only for the regression for the purpose of prediction, not for causal effect estimation.  
2. If the variable is not statistically significant, it has no impact on the outcome and we can drop it from the regression equation. 
3. Holding all else constant, the estimates from more data would produce smaller confidence intervals. 
4. Overlapping confidence intervals for small and large values of *Beauty* mean that there is no significant difference between the teaching evaluations of male and female instructors.  

::: 
::: {.column width="50%"}

::: {.r-stack}
```{r}
#| echo: false
#| fig-asp: 1
m0 <- lm(eval ~ beauty + female + age + minority + nonenglish, data = evals)

# create a grid for all possible combinations 
# tibble() only allows for different values in one variable 
plt <- expand_grid(
  beauty = 1:10,
  female = 0:1,
  age = mean(evals$age, na.rm = T),
  minority = 0,
  nonenglish = 0
) 

predict(m0,
        interval = "confidence",
        newdata = plt,
        level = 0.95) %>%
  bind_cols(plt) %>%
    mutate(female = if_else(female == 1, "Female", "Male")) %>%
  ggplot(aes(beauty, y = fit, 
             color = as_factor(female),
             fill  = as_factor(female))) +
  labs(
    x = "Beauty",
    y = "Average Teaching Evaluation",
    title = "Relationship between Instructor Beauty\nand Course Evaluations",
    # subtitle = "Intercept Shift with Same Slope",
    color = "Instructor is",
    fill = "Instructor is"
  ) +
  geom_line() +
  geom_ribbon(aes(ymin = lwr,
                  ymax = upr), 
                ) +
  # geom_pointrange(aes(y = fit, ymin = lwr, ymax = upr)) + # plots the parallel lines
  scale_color_viridis_d(end = 0.8, alpha = 0.8) +
  scale_fill_viridis_d(end = 0.8, alpha = 0.5) +
  theme(legend.position = "top") +
  scale_x_continuous(limits = c(1, 10), breaks = 1:10) +
  ylim(c(1, 5)) +
  theme(panel.grid.minor = element_blank(),
        plot.title.position = "plot",
        text = element_text(size = 20))
  #  geom_rug(
  #    data = augment(m0) %>%
  #   mutate(female = if_else(female == 1, "Female", "Male")),
  #    aes(x = beauty, y = eval),
  #   alpha = 0.3,
  #   sides = "b",
  #   position = "jitter",
  #   # size = 4,
  #   length = unit(0.05, "npc")
  # )
```

::: {.fragment}

```{r}
#| echo: false
#| fig-asp: 1
predict(m0,
        interval = "confidence",
        newdata = plt,
        level = 0.95) %>%
  bind_cols(plt) %>%
    mutate(female = if_else(female == 1, "Female", "Male")) %>%
  ggplot(aes(beauty, y = fit, 
             color = as_factor(female),
             fill  = as_factor(female))) +
  labs(
    x = "Beauty",
    y = "Average Teaching Evaluation",
    title = "Relationship between Instructor Beauty\nand Course Evaluations",
    # subtitle = "Intercept Shift with Same Slope",
    color = "Instructor is",
    fill = "Instructor is"
  ) +
  geom_line() +
  geom_ribbon(aes(ymin = lwr,
                  ymax = upr), 
                ) +
  # geom_pointrange(aes(y = fit, ymin = lwr, ymax = upr)) + # plots the parallel lines
  scale_color_viridis_d(end = 0.8, alpha = 0.8) +
  scale_fill_viridis_d(end = 0.8, alpha = 0.5) +
  theme(legend.position = "top") +
  scale_x_continuous(limits = c(1, 10), breaks = 1:10) +
  ylim(c(1, 5)) +
  theme(panel.grid.minor = element_blank(),
        plot.title.position = "plot",
        text = element_text(size = 20)) +
   geom_rug(
     data = augment(m0) %>%
    mutate(female = if_else(female == 1, "Female", "Male")),
     aes(x = beauty, y = eval),
    alpha = 0.3,
    sides = "b",
    position = "jitter",
    # size = 4,
    length = unit(0.05, "npc")
  )
```
:::
:::
::: 
:::


# Heterogeneous Effects 

## Which of These Hypotheses Specify Conditional Relationsip? {.smaller}

1. Women representatives are more likely than men representatives to focus on issues of relative importance to women.
2. The stronger the partisan identification of the voter, the stronger is anger in response to threat of electoral loss.
3. For candidates that the voter has seen on posters, appearance has a stronger effect on vote choice than for candidates that the voter has not seen on posters.
4. Women are more likely to substantively represent women if they were elected in districts with larger magnitude.
5. Government parties are more likely to engage in constructive legislative review than opposition parties.


## Dataset: Student Evaluations 

```{r}
glimpse(evals)
```

## Baseline Model {.smaller}

```{r}
#| echo: false
m0 <- lm(eval ~ beauty + female + age + nonenglish, 
         data = evals)
tidy(m0, conf.int = T) %>% 
  knitr::kable(digits = 3)
```

## Baseline Model {.smaller}
```{r}
#| echo: false
#| out-height: 50%
m0_df <- augment(m0, interval = "confidence")

# create a grid for all possible combinations 
# tibble() only allows for different values in one variable 
plt <- expand_grid(
  beauty = 0:10,
  female = 0:1,
  age = mean(m0_df$age),
  minority = 0,
  nonenglish = 0
) 
# 
# predict(m0,
#         interval = "confidence",
#         newdata = plt,
#         level = 0.95) %>%
#   bind_cols(plt) %>%
#     mutate(female = if_else(female == 1, "Female", "Male")) %>%
#   ggplot(aes(beauty, y = fit, 
#              color = as_factor(female),
#              fill  = as_factor(female))) +
#   labs(
#     x = "Beauty",
#     y = "Average Teaching Evaluation",
#     title = "Relationship between Instructor Beauty and Course Evaluations",
#     subtitle = "Intercept Shift with Same Slope",
#     color = "Instructor is",
#     fill = "Instructor is"
#   ) +
#   geom_line() +
#   # geom_ribbon(aes(ymin = lwr,
#   #                 ymax = upr)) +
#   # geom_pointrange(aes(y = fit, ymin = lwr, ymax = upr)) + # plots the parallel lines
#   scale_color_viridis_d(end = 0.8) +
#   scale_fill_viridis_d(end = 0.8, alpha = 0.5) +
#   theme(legend.position = "top") +
#   scale_x_continuous(limits = c(0, 10), breaks = 0:10) +
#   ylim(c(3.5, 4.8)) +
#   theme(panel.grid.minor = element_blank(),
#         plot.title.position = "plot")
plt <- expand_grid(
  beauty = 1:10,
  female = 0:1,
  age = mean(evals$age, na.rm = T),
  minority = 0,
  nonenglish = 0
) 
m0 <- lm(eval ~ beauty + female, data = evals)

vidata1 <- data.frame(beauty = 6, female = 0)
vidata2 <- data.frame(beauty = 7, female = 0)
gredata1<- data.frame(beauty = 3,female = 1)
gredata2<- data.frame(beauty = 4,female = 1)


predict(m0,
        interval = "confidence",
        newdata = plt,
        level = 0.95) %>%
  bind_cols(plt) %>%
    mutate(female = if_else(female == 1, "Female", "Male")) %>%
   ggplot(aes(beauty, y = fit,
             color = as_factor(female),
             fill  = as_factor(female))) +
  labs(
    x = "Beauty",
    y = "Average Teaching Evaluation",
    title = "Relationship between Instructor Beauty and Course Evaluations",
    subtitle = "Intercept Shift with Same Slope",
    color = "Instructor is",
    fill = "Instructor is"
  ) +
  annotate(
    "text",
    x = 7.25,
    y = 4.2,
    label = expression("β"[1]),
    size = 5
  ) +
  annotate(
    "point",
    x = 7,
    y = predict(m0, newdata = vidata2),
    size = 1.5,
    color = viridis::viridis(2)[1]
  ) +
  geom_segment(
    aes(
      x = 3,
      y = 3.5,
      xend = 4,
      yend = 3.5
    ),
    col = "black",
    arrow = arrow(
      length = unit(0.15, "cm"),
      ends = "both",
      type = "closed"
    )
  ) +
  geom_segment(
    aes(
      x = 7,
      y = 3.5,
      xend = 6,
      yend = 3.5
    ),
    col = "black",
    arrow = arrow(
      length = unit(0.15, "cm"),
      ends = "both",
      type = "closed"
    )
  ) +
  geom_segment(
    aes(
      x = 2,
      predict(m0, newdata = gredata2),
      xend = 2,
      yend =  predict(m0, newdata = gredata1)
    ),
    col = "black",
    arrow = arrow(
      length = unit(0.1, "cm"),
      ends = "both",
      type = "closed"
    )
  ) +
  geom_segment(
    aes(
      x = 2,
      y =  predict(m0, newdata = vidata1),
      xend = 2,
      yend =  predict(m0, newdata = vidata2)
    ),
    col = "black",
    arrow = arrow(
      length = unit(0.1, "cm"),
      ends = "both",
      type = "closed"
    )
  ) +
   # geom_segment(aes(x = 2, y =  predict(m0, newdata = vidata2), xend = 2, yend =  predict(m0, newdata = vidata1)), col = "black",
   #                arrow = arrow(length = unit(0.1, "cm"))) +
 geom_segment(aes(x= 4,xend = 4,y= 3,yend = predict(m0, newdata = gredata2)), color = "black", linetype = "dashed", size = 0.2) +
 annotate("point", x = 6, y = predict(m0, newdata = vidata1), size = 1.5, color = "darkviolet") +
 geom_segment(aes(x= 3,xend =3,y= 3,yend = predict(m0, newdata = gredata1)), color = "black", linetype = "dashed", size = 0.2) +
   geom_segment(aes(x= 6,xend =6,y= 3,yend = predict(m0, newdata = vidata1)), color = "black", linetype = "dashed", size = 0.2) +
   geom_segment(aes(x= 7,xend =7,y= 3,yend = predict(m0, newdata = vidata2)), color = "black", linetype = "dashed", size = 0.2) +
 annotate("text", x = 4.25, y = 3.8, label = expression("β"[1]), size = 5)+
 annotate("point", x = 3, y = predict(m0, newdata = gredata1), size = 1.5, color = "darkgreen") +
  annotate("point", x = 4, y = predict(m0, newdata = gredata2), size = 1.5, color = "darkgreen") +
geom_segment(aes(x= 1,xend = 7,y= predict(m0, newdata = vidata1),yend= predict(m0, newdata = vidata1)), color = "black",linetype = "dashed", size = 0.2) +
geom_segment(aes(x= 1,xend = 7,y= predict(m0, newdata = vidata2),yend= predict(m0, newdata = vidata2)), color = "black",linetype = "dashed", size = 0.2) +
geom_segment(aes(x= 1,xend = 4,y= predict(m0, newdata = gredata1),yend= predict(m0, newdata = gredata1)), color = "black",linetype = "dashed", size = 0.2) +
geom_segment(aes(x= 1,xend = 4,y= predict(m0, newdata = gredata2),yend= predict(m0, newdata = gredata2)), color = "black",linetype = "dashed", size = 0.2) +
  geom_line() +
  # geom_ribbon(aes(ymin = lwr,
  #                 ymax = upr)) +
  # geom_pointrange(aes(y = fit, ymin = lwr, ymax = upr)) + # plots the parallel lines
  scale_color_viridis_d(end = 0.8) +
  scale_fill_viridis_d(end = 0.8, alpha = 0.5) +
  theme(legend.position = "top") +
  scale_x_continuous(limits = c(1, 10), breaks = 1:10) +
  ylim(c(3, 4.5)) +
  theme(panel.grid.minor = element_blank(),
        plot.title.position = "plot")
```

## Heterogenous Effects of Beauty {.smaller}

<!-- - The effect of beauty on course evaluations differs across male and female professors: -->

```{r}
#| echo: false
m1 <- lm(eval ~ beauty * female + age +  nonenglish, data = evals)
tidy(m1, conf.int = T) %>% 
  knitr::kable(digits = 3)
```
 
## Heterogenous Effects of Beauty {.smaller}

```{r}
#| echo: false
#| out-height: 50%

m0 <- lm(eval ~ beauty * female, data = evals)

vidata1 <- data.frame(beauty = 6, female = 0)
vidata2 <- data.frame(beauty = 7, female = 0)
gredata1<- data.frame(beauty = 3,female = 1)
gredata2<- data.frame(beauty = 4,female = 1)


predict(m0,
        interval = "confidence",
        newdata = plt,
        level = 0.95) %>%
  bind_cols(plt) %>%
    mutate(female = if_else(female == 1, "Female", "Male")) %>%
   ggplot(aes(beauty, y = fit,
             color = as_factor(female),
             fill  = as_factor(female))) +
  labs(
    x = "Beauty",
    y = "Average Teaching Evaluation",
    title = "Relationship between Instructor Beauty and Course Evaluations",
    subtitle = "Intercept Shift with Same Slope",
    color = "Instructor is",
    fill = "Instructor is"
  ) +
  annotate(
    "text",
    x = 8.5,
    y = 4.3,
    label = expression("β"[2] + paste("β"[3], "Beauty")),
    size = 5
  ) +
  annotate(
    "point",
    x = 7,
    y = predict(m0, newdata = vidata2),
    size = 1.5,
    color = viridis::viridis(2)[1]
  ) +
  geom_segment(
    aes(
      x = 3,
      y = 3.5,
      xend = 4,
      yend = 3.5
    ),
    col = "black",
    arrow = arrow(
      length = unit(0.15, "cm"),
      ends = "both",
      type = "closed"
    )
  ) +
  geom_segment(
    aes(
      x = 7,
      y = 3.5,
      xend = 6,
      yend = 3.5
    ),
    col = "black",
    arrow = arrow(
      length = unit(0.15, "cm"),
      ends = "both",
      type = "closed"
    )
  ) +
  geom_segment(
    aes(
      x = 2,
      predict(m0, newdata = gredata2),
      xend = 2,
      yend =  predict(m0, newdata = gredata1)
    ),
    col = "black",
    arrow = arrow(
      length = unit(0.1, "cm"),
      ends = "both",
      type = "closed"
    )
  ) +
  geom_segment(
    aes(
      x = 2,
      y =  predict(m0, newdata = vidata1),
      xend = 2,
      yend =  predict(m0, newdata = vidata2)
    ),
    col = "black",
    arrow = arrow(
      length = unit(0.1, "cm"),
      ends = "both",
      type = "closed"
    )
  ) +
   # geom_segment(aes(x = 2, y =  predict(m0, newdata = vidata2), xend = 2, yend =  predict(m0, newdata = vidata1)), col = "black",
   #                arrow = arrow(length = unit(0.1, "cm"))) +
 geom_segment(aes(x= 4,xend = 4,y= 3,yend = predict(m0, newdata = gredata2)), color = "black", linetype = "dashed", size = 0.2) +
 annotate("point", x = 6, y = predict(m0, newdata = vidata1), size = 1.5, color = "darkviolet") +
 geom_segment(aes(x= 3,xend =3,y= 3,yend = predict(m0, newdata = gredata1)), color = "black", linetype = "dashed", size = 0.2) +
   geom_segment(aes(x= 6,xend =6,y= 3,yend = predict(m0, newdata = vidata1)), color = "black", linetype = "dashed", size = 0.2) +
   geom_segment(aes(x= 7,xend =7,y= 3,yend = predict(m0, newdata = vidata2)), color = "black", linetype = "dashed", size = 0.2) +
 annotate("text", x = 4.25, y = 3.8, label = expression("β"[2]), size = 5)+
 annotate("point", x = 3, y = predict(m0, newdata = gredata1), size = 1.5, color = "darkgreen") +
  annotate("point", x = 4, y = predict(m0, newdata = gredata2), size = 1.5, color = "darkgreen") +
geom_segment(aes(x= 1,xend = 7,y= predict(m0, newdata = vidata1),yend= predict(m0, newdata = vidata1)), color = "black",linetype = "dashed", size = 0.2) +
geom_segment(aes(x= 1,xend = 7,y= predict(m0, newdata = vidata2),yend= predict(m0, newdata = vidata2)), color = "black",linetype = "dashed", size = 0.2) +
geom_segment(aes(x= 1,xend = 4,y= predict(m0, newdata = gredata1),yend= predict(m0, newdata = gredata1)), color = "black",linetype = "dashed", size = 0.2) +
geom_segment(aes(x= 1,xend = 4,y= predict(m0, newdata = gredata2),yend= predict(m0, newdata = gredata2)), color = "black",linetype = "dashed", size = 0.2) +
  geom_line() +
  # geom_ribbon(aes(ymin = lwr,
  #                 ymax = upr)) +
  # geom_pointrange(aes(y = fit, ymin = lwr, ymax = upr)) + # plots the parallel lines
  scale_color_viridis_d(end = 0.8) +
  scale_fill_viridis_d(end = 0.8, alpha = 0.5) +
  theme(legend.position = "top") +
  scale_x_continuous(limits = c(1, 10), breaks = 1:10) +
  ylim(c(3, 4.5)) +
  theme(panel.grid.minor = element_blank(),
        plot.title.position = "plot")
# predict(m1,
#         interval = "confidence",
#         newdata = plt,
#         level = 0.95) %>%
#   bind_cols(plt) %>%
#     mutate(female = if_else(female == 1, "Female", "Male")) %>%
#   ggplot(aes(beauty, y = fit, 
#              color = as_factor(female),
#              fill  = as_factor(female))) +
#   labs(
#     x = "Beauty",
#     y = "Average Teaching Evaluation",
#     title = "Relationship between Instructor Beauty and Course Evaluations",
#     subtitle = "Intercept Shift with Different Slopes",
#     color = "Instructor is",
#     fill = "Instructor is"
#   ) +
#   geom_line() +
#   # geom_ribbon(aes(ymin = lwr,
#   #                 ymax = upr)) +
#   # geom_pointrange(aes(y = fit, ymin = lwr, ymax = upr)) + # plots the parallel lines
#   scale_color_viridis_d(end = 0.8) +
#   scale_fill_viridis_d(end = 0.8, alpha = 0.5) +
#   theme(legend.position = "top") +
#   scale_x_continuous(limits = c(0, 10), breaks = 0:10) +
#   ylim(c(3, 5)) +
#   theme(panel.grid.minor = element_blank(),
#         plot.title.position = "plot")
# # m0_df <- augment(m0, interval = "confidence")

```


## Heterogenous Effects: Model Specification {.smaller}

$$
\text{Evaluation}_i = \underbrace{\hat\beta_0}_{\text{Expected evaluation}\\\text{for males if}\\\text{beauty score = 0}} + \underbrace{\hat\beta_1}_{\text{Difference in}\\\text{evaluation between}\\\text{males and females}\\\text{if beauty score = 0}} \text{Female}_i + \underbrace{\hat\beta_2}_{\text{Effect of each}\\\text{additional point}\\\text{in score for males}} \text{Beauty}_i \\ +  \underbrace{\hat\beta_3}_{\text{Difference in}\\\text{effect of beauty}\\\text{between males}\\\text{and females}} \text{Female}_i \times \text{Beauty}_i
$$

- $\text{Female}_i$ and $\text{Beauty}_i$: constitutive terms 
- $\text{Female}_i \times \text{Beauty}_i$: interaction term 
 
## Heterogenous Effects: Model Specification {.smaller}
 
We are allowing the effects of beauty to vary across the subgroups:

Male instructors:

$$
\text{Evaluation}_i = \hat\beta_0 + \hat\beta_1 \times [\text{Female}_i = 0] + \hat\beta_2 \text{Beauty}_i +  \hat\beta_3 \times [\text{Female}_i = 0] \times \text{Beauty}_i \\
\text{Evaluation}_i = \hat\beta_0 + \hat\beta_1 \times 0 + \hat\beta_2 \text{Beauty}_i +  \hat\beta_3 \times 0 \times \text{Beauty}_i \\
\text{Evaluation}_i = {\hat\beta_0} + {\hat\beta_2} \text{Beauty}_i
$$

Female instructors:

$$
\text{Evaluation}_i = \hat\beta_0 + \hat\beta_1 \times [\text{Female}_i = 1] + \hat\beta_2 \text{Beauty}_i +  \hat\beta_3 \times [\text{Female}_i = 1] \times \text{Beauty}_i \\
\text{Evaluation}_i = (\hat\beta_0 + \hat\beta_1)  + (\hat\beta_2  +  \hat\beta_3 ) \times \text{Beauty}_i 
$$

## Working with Interaction Effects {.smaller}

```{r}
#| results: asis
#| echo: fenced
m1 <- lm(eval ~ beauty * female + age + nonenglish, data = evals)
equatiomatic::extract_eq(m1, use_coefs = T, terms_per_line = 3, coef_digits = 3)
```

- Include all constitutive terms (i.e. those variables that compose the interaction)
- Do not interpret the coefficients on constitutive terms as unconditional marginal effects:

> Wrong: On average, beauty score is associated with an increase in teaching evaluations, holding all else constant. \
> Wrong: On average, female instructors receive 0.05 points more in teaching evaluations than males, holding all else constant. \
> Right: On average, beauty tends to have smaller effect on teaching evaluations among female instructors in comparison to male instructors, holding all else constant.  

## Causal Interaction vs. Effect Modification {.smaller}

$$
\text{Evaluation}_i = \hat\beta_0 + \hat\beta_1 \times \text{Female}_i  + \hat\beta_2 \text{Beauty}_i +  \hat\beta_3 \times \text{Female}_i \times \text{Beauty}_i \\
$$

Mathematically, interactions allow us to tell two stories: 

- the effect of *Beauty* on *Teaching Evaluations* differs between male and female instructors 
- the effect of *Gender* on *Teaching Evaluations* differs across values of beauty score  

:::task 
Interpret the interaction coefficients in line with your causal story and your selected control variables.
:::


## Alternative: Using Subsets {.smaller}

::: columns
::: {.column width="50%"}

```{r}
m3 <- lm(eval ~ beauty + age + nonenglish, 
         data = evals %>% filter(female == 1))
tidy(m3, conf.int = T, conf.level = 0.9) %>% 
  knitr::kable(digits = 3)
```

::: 
::: {.column width="50%"}

```{r}
m4 <- lm(eval ~ beauty + age + nonenglish, 
         data = evals %>% filter(female == 0))
tidy(m4, conf.int = T, conf.level = 0.9) %>% 
  knitr::kable(digits = 3)
```
::: 
::: 

- Sample size: interactions rely on more data, hence smaller uncertainty 
- Comparisons are easier across effects with one model 
- Works straightforwardly when the moderator is categorical, continuous moderators require arbitrary decisions about cutoff points 
- Best: do both to make sure your results are robust 


## Intractions with Continous Variables {.smaller}

$$
\text{Evaluation}_i = \hat\beta_0 + \hat\beta_1 \times \text{Age}_i  + \hat\beta_2 \text{Beauty}_i +  \hat\beta_3 \times \text{Age}_i \times \text{Beauty}_i \\
$$

- Marginal effect of *Beauty* on *Teaching evaluation*: $\hat\beta_2  + \hat\beta_3 \times \text{Age}_i$

- Marginal effect of *Age* on *Teaching evaluation*: $\hat\beta_1  +  \hat\beta_3 \times \text{Beauty}_i$

## Quiz: What is the Marginal Effect of *Beauty*?

$$
\text{Evaluation}_i = \hat\beta_0 + \hat\beta_1 \times \text{Age}_i  + \hat\beta_2 \text{Beauty}_i +  \hat\beta_3 \times \text{Age}_i \times \text{Beauty}_i \\
+ \hat\beta_4 \text{Female}_i +  \hat\beta_5 \times \text{Female}_i \times \text{Beauty}_i
$$

1. $\hat\beta_2  +  \hat\beta_3 \times \text{Age}_i$
2. $\hat\beta_2  +  \hat\beta_3 \times \text{Age}_i + \hat\beta_5 \times \text{Female}_i$
3. $\hat\beta_2  +  \hat\beta_5 \times \text{Female}_i$
4. $\hat\beta_3 \times \text{Age}_i + \hat\beta_5 \times \text{Female}_i$

## Plotting Interactions with Continuous Variables {.smaller}

```{r}
#| echo: false
m2 <- lm(eval ~ beauty * age + female  + nonenglish, data = evals)
# tidy(m2, conf.int = T, conf.level = 0.9) %>% 
#   knitr::kable(digits = 3)
```


```{r}
#| echo: false
plt <- expand_grid(
  beauty = 0:10,
  female = 0,
  age = quantile(evals$age, probs = c(0.1, 0.5, 0.9)),
  minority = 0,
  nonenglish = 0
) 

predict(m2,
        interval = "confidence",
        newdata = plt,
        level = 0.95) %>%
  bind_cols(plt) %>%
  ggplot(aes(beauty, y = fit, 
             color = as_factor(age),
             fill  = as_factor(age))) +
  labs(
    x = "Beauty",
    y = "Average Teaching Evaluation",
    title = "Relationship between Instructor Beauty and Course Evaluations",
    # subtitle = "Intercept Shift with Different Slopes",
    color = "Instructor's age (in years)",
    fill = "Instructor's age  (in years)"
  ) +
  geom_line() +
  # geom_ribbon(aes(ymin = lwr,
  #                 ymax = upr)) +
  # geom_pointrange(aes(y = fit, ymin = lwr, ymax = upr)) + # plots the parallel lines
  scale_color_viridis_d(end = 0.8) +
  scale_fill_viridis_d(end = 0.8, alpha = 0.5) +
  theme(legend.position = "top") +
  scale_x_continuous(limits = c(0, 10), breaks = 0:10) +
  ylim(c(3, 5)) +
  theme(panel.grid.minor = element_blank(),
        plot.title.position = "plot") 


```

## Plotting Interactions with Continuous Variables 

```{r}
#| layout-ncol: 2
#| echo: false
#| fig-asp: 1
predict(m2,
        interval = "confidence",
        newdata = plt,
        level = 0.95) %>%
  bind_cols(plt) %>%
  ggplot(aes(beauty, y = fit, 
             color = as_factor(age),
             fill  = as_factor(age))) +
  labs(
    x = "Beauty",
    y = "Average Teaching Evaluation",
    title = "Relationship between Instructor Beauty and Course Evaluations",
    # subtitle = "Intercept Shift with Different Slopes",
    color = "Instructor's age (in years)",
    fill = "Instructor's age  (in years)"
  ) +
  geom_line() +
  # geom_ribbon(aes(ymin = lwr,
  #                 ymax = upr)) +
  # geom_pointrange(aes(y = fit, ymin = lwr, ymax = upr)) + # plots the parallel lines
  scale_color_viridis_d(end = 0.8) +
  scale_fill_viridis_d(end = 0.8, alpha = 0.5) +
  theme(legend.position = "top") +
  scale_x_continuous(limits = c(0, 10), breaks = 0:10) +
  ylim(c(3, 5)) +
  theme(panel.grid.minor = element_blank(),
        plot.title.position = "plot") 


plt <- expand_grid(
  beauty = c(2, 4, 6, 8),
  female = 0,
  age = seq(min(evals$age, na.rm = T), max(evals$age, na.rm = T), length.out = 10),
  minority = 0,
  nonenglish = 0
) 

predict(m2,
        interval = "confidence",
        newdata = plt,
        level = 0.95) %>%
  bind_cols(plt) %>%
  ggplot(aes(age, y = fit, 
             color = as_factor(beauty),
             fill  = as_factor(beauty))) +
  labs(
    x = "Instructor's age (in years)",
    y = "Average Teaching Evaluation",
    title = "Relationship between Instructor Beauty and Course Evaluations",
    # subtitle = "Intercept Shift with Different Slopes",
    color = "Beauty Score",
    fill = "Beauty Score"
  ) +
  geom_line() +
  # geom_ribbon(aes(ymin = lwr,
  #                 ymax = upr)) +
  # geom_pointrange(aes(y = fit, ymin = lwr, ymax = upr)) + # plots the parallel lines
  scale_color_viridis_d(end = 0.8) +
  scale_fill_viridis_d(end = 0.8, alpha = 0.5) +
  theme(legend.position = "top") +
  # scale_x_continuous(limits = c(0, 10), breaks = 0:10) +
  ylim(c(3, 5)) +
  theme(panel.grid.minor = element_blank(),
        plot.title.position = "plot") 
```

## Plotting Interactions with Continuous Variables II

```{r}
#| layout-ncol: 2
#| echo: false
#| fig-asp: 1

plt <- expand_grid(
  beauty = 0:10,
  female = 0,
  age = quantile(evals$age, probs = c(0.1, 0.5, 0.9)),
  minority = 0,
  nonenglish = 0
)

predict(m2,
        interval = "confidence",
        newdata = plt,
        level = 0.95) %>%
  bind_cols(plt) %>%
  ggplot(aes(beauty, y = fit, 
             color = as_factor(age),
             fill  = as_factor(age))) +
  labs(
    x = "Beauty",
    y = "Average Teaching Evaluation",
    title = "Relationship between Instructor Beauty and Course Evaluations",
    # subtitle = "Intercept Shift with Different Slopes",
    color = "Instructor's age (in years)",
    fill = "Instructor's age  (in years)"
  ) +
  geom_line() +
  geom_ribbon(aes(ymin = lwr,
                  ymax = upr)) +
  # geom_pointrange(aes(y = fit, ymin = lwr, ymax = upr)) + # plots the parallel lines
  scale_color_viridis_d(end = 0.8) +
  scale_fill_viridis_d(end = 0.8, alpha = 0.5) +
  theme(legend.position = "top") +
  scale_x_continuous(limits = c(0, 10), breaks = 0:10) +
  ylim(c(3, 5.5)) +
  theme(panel.grid.minor = element_blank(),
        plot.title.position = "plot") 


plt <- expand_grid(
  beauty = c(2, 4, 6, 8),
  female = 0,
  age = seq(min(evals$age, na.rm = T), max(evals$age, na.rm = T), length.out = 10),
  minority = 0,
  nonenglish = 0
) 

predict(m2,
        interval = "confidence",
        newdata = plt,
        level = 0.95) %>%
  bind_cols(plt) %>%
  ggplot(aes(age, y = fit, 
             color = as_factor(beauty),
             fill  = as_factor(beauty))) +
  labs(
    x = "Instructor's age (in years)",
    y = "Average Teaching Evaluation",
    title = "Relationship between Instructor Beauty and Course Evaluations",
    # subtitle = "Intercept Shift with Different Slopes",
    color = "Beauty Score",
    fill = "Beauty Score"
  ) +
  geom_line() +
  geom_ribbon(aes(ymin = lwr,
                  ymax = upr)) +
  # geom_pointrange(aes(y = fit, ymin = lwr, ymax = upr)) + # plots the parallel lines
  scale_color_viridis_d(end = 0.8) +
  scale_fill_viridis_d(end = 0.8, alpha = 0.5) +
  theme(legend.position = "top") +
  # scale_x_continuous(limits = c(0, 10), breaks = 0:10) +
  ylim(c(3, 5.5)) +
  theme(panel.grid.minor = element_blank(),
        plot.title.position = "plot") 
```


# Appendix: Using Simulations for Interpretation 

## Approaches to Obtain Confidence Intervals {.smaller}

- So far, we used stats to calculate confidence intervals: 
  - assumed sampling distributions are normal 
  - calculated point estimates and standard errors
  - calculate confidence intervals from estimated SEs and critical values for $\mathcal{N}(0, 1)$
  
- Another approach utilizes simulations:
  - assumed sampling distributions are normal 
  - calculated point estimates and standard errors
  - use these values to obtain sets of simulated coefficients
  - use these simulated coefficients to make predictions 

## Draws from Simulated Sampling Distributions {.smaller}

We would use our coefficient estimates and uncertainty about them to calculate sampling distributions (our best guesses about them based on our data)

```{r}
#| echo: false
library(clarify)
set.seed(2023)
m1 <- lm(eval ~ beauty * female, data = evals)

# get draws from multivariate normal distribution 
sims <- clarify::sim(m1)
as_tibble(sims$sim.coefs) %>%
  clean_names() %>%
  pivot_longer(values_to = "estimate",
               names_to = "term",
               cols = everything()) %>%
  ggplot(aes(x = estimate)) +
  geom_histogram(color = "white")   +
  facet_wrap(~term, scales = "free") +
  labs(title = "Simulated Coefficients")
```


## Calculating Expected Values {.smaller}

$${ \begin{array}{c}
        \tilde{\beta}_0^1 \times 1+\tilde{\beta}_1^1 \times   \text{Age}_i  +\tilde{\beta}_2^1 \times \text{Beauty}_i + \tilde{\beta}_3^1 \times \text{Age}_i \times \text{Beauty}_i &= \tilde{y}^1\\
 \tilde{\beta}_0^2 \times 1+\tilde{\beta}_1^2 \times   \text{Age}_i  +\tilde{\beta}_2^2 \times \text{Beauty}_i + \tilde{\beta}_3^2 \times \text{Age}_i \times \text{Beauty}_i &= \tilde{y}^2\\
  \tilde{\beta}_0^3 \times 1+\tilde{\beta}_1^3 \times   \text{Age}_i  +\tilde{\beta}_2^3 \times \text{Beauty}_i + \tilde{\beta}_3^3 \times \text{Age}_i \times \text{Beauty}_i &= \tilde{y}^3\\
        \dots \\
        \tilde{\beta}_0^{1000} \times 1+\tilde{\beta}_1^{1000} \times \text{Age}_i  +\tilde{\beta}_2^{1000} \times \text{Beauty}_i + \tilde{\beta}_3^{1000} \times \text{Age}_i \times \text{Beauty}_i &= \tilde{y}^{1000}\\
    \end{array} }$$


## Plotting Expected Values {.smaller}

```{r}
#| echo: false
# obtain predictions based on specified values
results <- sim_setx(sims, 
                    x = list(female = 0:1,
                             beauty = 1:10))

plot(results) +
 scale_color_viridis_d(end = 0.8) +
  scale_fill_viridis_d(end = 0.8, alpha = 0.5) +
  theme(legend.position = "top") +
  scale_x_continuous(limits = c(1, 10), breaks = 1:10) +
  ylim(c(3, 5)) +
  theme(panel.grid.minor = element_blank(),
        plot.title.position = "plot") +
  labs(y = "Average Teaching Evaluation")

```

## Evaluating the Effects with First Differences

```{r}
#| echo: false
FD <- sim_setx(sims,
               x = list(female = 0,
                        beauty = 1:10)) -
  sim_setx(sims,
           x = list(female = 1,
                    beauty = 1:10))

plot(FD) +
 scale_color_viridis_d(end = 0.8) +
  scale_fill_viridis_d(end = 0.8, alpha = 0.5) +
  theme(legend.position = "top") +
  scale_x_continuous(limits = c(1, 10), breaks = 1:10) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  # ylim(c(3, 5)) +
  theme(panel.grid.minor = element_blank(),
        plot.title.position = "plot") +
  labs(y = "Difference in Average Teaching Evaluation",
       subtitle = "First Difference between Male and Female Instructors")
```




## Using Simulations for Interpretation

```{r}
#| eval: false
library(clarify)
set.seed(2023)
m1 <- lm(eval ~ beauty * female, data = evals)

# get draws from multivariate normal distribution 
sims <- clarify::sim(m1)

# obtain predictions based on specified values
results <- sim_setx(sims, 
                    x = list(female = 0:1,
                             beauty = 1:10))

plot(results) 

FD <- sim_setx(sims,
               x = list(female = 0,
                        beauty = 1:10)) -
  sim_setx(sims,
           x = list(female = 1,
                    beauty = 1:10))

plot(FD) +
 scale_color_viridis_d(end = 0.8) +
  scale_fill_viridis_d(end = 0.8, alpha = 0.5) +
  theme(legend.position = "top") +
  scale_x_continuous(limits = c(1, 10), breaks = 1:10) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  # ylim(c(3, 5)) +
  theme(panel.grid.minor = element_blank(),
        plot.title.position = "plot") +
  labs(y = "Difference in Average Teaching Evaluation",
       subtitle = "First Difference between Male and Female Instructors")

plot(FD) +
 scale_color_viridis_d(end = 0.8) +
  scale_fill_viridis_d(end = 0.8, alpha = 0.5) +
  theme(legend.position = "top") +
  scale_x_continuous(limits = c(1, 10), breaks = 1:10) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  # ylim(c(3, 5)) +
  theme(panel.grid.minor = element_blank(),
        plot.title.position = "plot") +
  labs(y = "Difference in Average Teaching Evaluation",
       subtitle = "First Difference between Male and Female Instructors")
```


<!-- # Intro -->


<!-- ```{r} -->
<!-- #| echo: false -->
<!-- countdown(minutes = 4, color_background = "white", left = 1) -->
<!-- ``` -->



<!-- ::: columns -->
<!-- ::: {.column width="30%"} -->

<!-- ```{dot} -->
<!-- //| echo: false -->
<!-- //| fig-width: 3 -->
<!-- //| fig-height: 3 -->
<!-- //| out-width: 100% -->
<!-- digraph D { -->
<!--   node [shape=oval, fontname="AtkinsonHyperlegible-Regular"]; -->
<!--   edge [minlen = 1.2, arrowhead = vee, arrowsize=0.5]; -->
<!--   a [label = "Beauty Score"]; -->
<!--   b [label = "Evaluations"]; -->
<!--   c [label = "Female"]; -->

<!--   {rank=same a b}; -->
<!--   {rank=sink c}; -->
<!--   a->b; -->
<!--   c->a; -->
<!--   c->b; -->
<!-- } -->
<!-- ``` -->


<!-- :::  -->
<!-- ::: {.column width="70%"} -->

<!-- 1.  After specifying the DGP in a DAG and following our identification strategy, we can interpret the estimated coefficient for the effect of *Beauty Score* on *Evaluations* causally. -->

<!-- 2.  After specifying the DGP in a DAG and following our identification strategy, we can interpret the estimated coefficient for the effect of *Female* on *Evaluations* causally. -->

<!-- 3. On average, a 5-point increase in *Beauty Score* would be associated with 0.4-point increase in *Teaching Evaluations*, for both female and female instructors.  -->

<!-- 4. The effect of *Female* on *Teaching Evaluations* would be represented through the change in the slope of the regression coefficient for *Beauty Score*.  -->

<!-- 5. For $evals = \beta_0 + \beta_1 \cdot Female$,  $\beta_1$ would be the equal to difference-in-means estimator.  -->

<!-- :::  -->
<!-- :::  -->
